
## Thermodynamics
### Entropy
```{r entropyDef, echo=FALSE}
question("Which definition of entropy is correct?",
  answer("Entropy is the total heat content of a system"),
  answer("Entropy is the maximum amount of energy that is available to do work"),
  answer("Entropy is a measure for the degree of disorder", correct = TRUE),
  answer("Entropy is the free energy of each component of the system"),
  allow_retry = TRUE
)
```

```{r entropyChange, echo=FALSE}
question("Which of the following statements is correct?",
  answer("The entropy of the universe is constant"),
  answer("The entropy of the universe decreases over time"),
  answer("The entropy of the universe increases over time", correct = TRUE),
  allow_retry = TRUE
)
```


```{r mostProbableOutcome, echo=FALSE}
question_radio(
  "The most probable outcome is the macroscopic state with the greatest number of corresponding microscopic states",
  answer("TRUE", correct = TRUE),
  answer("FALSE"),
  allow_retry = TRUE
)
```

How many possible microstates are there when flipping 10 coins?

```{r microstates, exercise = TRUE}

```

```{r microstates-solution}
n <- 2 # number of options (heads or tails)
k <- 10 # number of items (number of coins)
microstates <- n^k
print(microstates)
```

```{r microstates-check}
grade_result(
  pass_if(~identical(.result, 1024))
)
```

The number of microstates for a particular macrostate is given as: $\binom{n}{x} = \frac{n!}{x!(n-x)!}$, where $n$ is the number of coins and $x$ is the number of times you receive heads. The term $\binom{n}{x}$ is read "$n$ choose $x$". Assuming you have 20 coins, how many microstates are there for receiving 10 times heads?

```{r numberOfMicrostatesHalf, exercise = TRUE, exercise.eval = FALSE}
n <- 20
x <- 10
factorial(___)
```

```{r numberOfMicrostatesHalf-solution}
n <- 20 # number of coins
x <- 10 # number of heads (or tails)
answer <- factorial(n) / (factorial(x) * factorial(n - x))
print(answer)
```

```{r numberOfMicrostatesHalf-check}
grade_result(
  pass_if(~identical(.result, 184756))
)
```

How many microstates are there for receiving 5 times heads when flipping 20 coins?

```{r numberOfMicrostatesQuarter, exercise = TRUE, exercise.eval = FALSE}

```

```{r numberOfMicrostatesQuarter-solution}
n <- 20 # number of coins
x <- 5 # number of heads (or tails)
answer <- factorial(n) / (factorial(x) * factorial(n - x))
print(answer)
```

```{r numberOfMicrostatesQuarter-check}
grade_result(
  pass_if(~identical(.result, 15504))
)
```

How much more likely is it to receive 10 compared to 5 times heads when flipping 20 coins?

```{r probability, exercise = TRUE, exercise.eval = FALSE}

```

```{r probability-solution}
answer <- 184756/15504
print(answer)
```

```{r probability-check}
grade_result(
  pass_if(~identical(.result, 184756/15504))
)
```

Consider a sample of ice with a mass of 3 kg that melts at a constant temperature of about 273.15K. Calculate the corresponding change in entropy. Assume that the energy required for melting (i.e. the heat of fusion) equals 3.34 x 10^5^ J/kg.

```{r meltingIce, exercise = TRUE, exercise.eval = FALSE}
T <- ___ # temperature (K)
ice.mass <- ___ # ice mass (kg)
Lf <- ___ # heat of fusion (J/kg)

# Heat added to melt ice
dQ <- ___

# Change in entropy
dS <- ___
print(dS)
```

```{r meltingIce-solution}
T <- 273.15 # temperature (K)
ice.mass <- 3 # ice mass (kg)
Lf <- 3.34 * 10^5 # heat of fusion (J/kg)

# Heat added to melt ice
dQ <- Lf * ice.mass # J

# Change in entropy
dS <- dQ / T # J/K
print(dS)
```

```{r meltingIce-check}
grade_result(
  pass_if(~identical(round(.result), round(3668.314)))
)
```

```{r sortUnits, echo = FALSE}

units <- c("W", "J", "N", "J", "Pa", "J")
rank_units <- sortable::question_rank(
  "Sort the units in the order of the following terms: power, energy, force, heat, pressure, work",
  learnr::answer(units, correct = TRUE),
  learnr::answer(rev(units), correct = FALSE, "Other direction!"),
  allow_retry = TRUE
)
rank_units
```

Consider a sample of water with a volume of 1 cm^3^ that is boiled at constant atmospheric surface pressure of 101300 Pa. The resulting steam occupies a volume of 1500 cm^3^. Calculate the change in internal energy. Assume that the latent heat of vaporization equals 2.256 x 10^6^ J kg^-1^.

```{r boilingWater, exercise = TRUE, exercise.eval = FALSE}
# Constants
Lv <- ___ # latent heat of vaporization (J/kg)
rho <- ___ # water density (kg/m3)

# Inputs
p <- ___ # pressure (Pa)
V1 <- ___ # volume (m3)
V2 <- ___ # volume (m3)

# Work done by the water when it vaporizes (J)
dW <- ___

# Water mass (kg)
m <- ___

# Heat added to the water (J)
dQ <- ___

# Change in internal energy (J)
dU <- ___
print(dU)
```

```{r boilingWater-solution}
# Constants
Lv <- 2.256 * 10^6 # latent heat of vaporization (J/kg)
rho <- 1000 # water density (kg/m3)

# Inputs
p <- 101300 # pressure (Pa)
V1 <- 1/1000000 # volume (m3)
V2 <- 1500/1000000 # volume (m3)

# Work done by the water when it vaporizes (J)
dW <- p * (V2 - V1)

# Water mass (kg)
m <- V1 * rho

# Heat added to the water (J)
dQ <- m * Lv

# Change in internal energy (J)
dU <- dQ - dW
print(dU)
```

```{r boilingWater-check}
grade_result(
  pass_if(~identical(round(.result), round(2104.151)))
)
```

### Gibbs free energy

```{r entropyVsGibbsQuestions, echo=FALSE}
question("Which of the following statements on Gibbs free energy and entropy are true? (select all that apply)",
  answer("Entropy is measured in J/K while Gibbs free energy is measured in J", correct = TRUE),
  answer("Gibbs free energy and entropy only differ in sign"),
  answer("The change in Gibbs free energy equals the change in enthalpy minus the change in entropy times temperature", correct = TRUE),
  answer("The change in Gibbs free energy equals the change in enthalpy plus the change in entropy divided by temperature"),
  answer("Any process occurs spontaneously if entropy increases"),
  answer("Any process occurs spontaneously if Gibbs free energy decreases", correct = TRUE),
  allow_retry = TRUE
)
```

Under standard conditions, the oxidation of nitric oxide to nitrogen dioxide (2NO + O$_2$ $\rightarrow$ 2NO$_2$) causes a change in enthalpy by -114100 J and a change in entropy by -146.5 J/K. Calculated the associated change in Gibbs free energy for a temperature of 298.15 K.

```{r nitricOxide, exercise = TRUE, exercise.eval = FALSE}

```

```{r nitricOxide-solution}
dH <- -114100 # change in enthalpy (J)
dS <- -146.5 # change in entropy (J/K)
T <- 298.15 # Temperature (K)

dG <- dH - T * dS
print(dG)
```

```{r nitricOxide-check}
grade_result(
  pass_if(~identical(round(.result), round(-70421.02)))
)
```

```{r Gibbs, echo=FALSE}
question("Check which of the statements below is correct. The oxidation of nitric oxide to nitrogen dioxide as described above occurs spontaneously because:",
  answer("the change in enthalpy is negative"),
  answer("the change in entropy is negative"),
  answer("the change in Gibbs free energy is negative", correct = TRUE),
  allow_retry = TRUE
)
```

```{r spontaneity, echo=FALSE}
question_radio(
  "How does the temperature need to change to make the reaction described above non-spontaneous?",
  answer("Temperature must increase", correct = TRUE),
  answer("Temperature must decreae"),
  allow_retry = TRUE
)
```

### Chemical potential
```{r GibbsAndChemPotQuestions, echo=FALSE}
question("Which of the following statements on Gibbs free energy and chemical potential is correct? (select all that apply)",
answer("Chemical potential gives the free energy of each chemical species of the system", correct = TRUE),
answer("The free energy of each species contributes to the Gibbs free energy of the system", correct = TRUE),
answer("Gibbs free energy gives the free energy of each chemical species of the system"),
  allow_retry = TRUE
)
```

The chemical potential can be defined as: $$\mu=\mu^{*}+RT \ln{c}+\bar{V}p+zF\phi_{E}+mgh$$
Create a Figure that shows how term II varies as a function of mole fraction (x-axis) for a range of temperatures (different lines). Allow the mole fraction to range between 0.1 and 1.0 (-) and temperature to range from 280 to 320 K.

```{r activityTerm, exercise = TRUE, exercise.eval = FALSE}
# constants
R <- 8.314 # universal gas constant (J K–1 mol–1)

# variables
c_min <- ___ # concentration of solutes (mol m–3)
c_max <- ___ # concentration of solutes (mol m–3)
c <- seq(c_min, c_max, 0.01)

T_min <- ___ # absolute temperature (K) 
T_max <- ___ # absolute temperature (K) 
T <- seq(T_min, T_max, 10) # temperature (K)

my.fun <- function(c, T){
  y <- ___
  print(y)
}

data <- outer(c, T, my.fun)

my.xlab <- "Mole fraction (-)"
my.ylab <- "Chemical Potential (J mol$^{-1}$)"

my.xlab <- latex2exp::TeX(my.xlab)
my.ylab <- latex2exp::TeX(my.ylab)

my.col <- viridis::plasma(n = ncol(data))

matplot(x = c, y = data, type = "l", lty = 1, col = my.col,
        xlab = my.xlab, ylab = my.ylab)

legend.text <- paste(T, "T", sep = " ")
legend("bottomright", legend = legend.text, pch = 16, col = my.col, bty = "n")

```

```{r activityTerm-solution}
# constants
R <- 8.314 # universal gas constant (J K–1 mol–1)

# variables
c_min <- 0.1 # concentration of solutes (mol m–3)
c_max <- 1 # concentration of solutes (mol m–3)
c <- seq(c_min, c_max, 0.01)

T_min <- 280 # absolute temperature (K) 
T_max <- 320 # absolute temperature (K) 
T <- seq(T_min, T_max, 10) # temperature (K)

my.fun <- function(c, T){
  y <- R * T * log(c)
  print(y)
}

data <- outer(c, T, my.fun)

my.xlab <- "Mole fraction (-)"
my.ylab <- "Chemical Potential (J mol$^{-1}$)"

my.xlab <- latex2exp::TeX(my.xlab)
my.ylab <- latex2exp::TeX(my.ylab)

my.col <- viridis::plasma(n = ncol(data))

matplot(x = c, y = data, type = "l", lty = 1, col = my.col,
        xlab = my.xlab, ylab = my.ylab)

legend.text <- paste(T, "T", sep = " ")
legend("bottomright", legend = legend.text, pch = 16, col = my.col, bty = "n")

```

```{r activityTermQuestions, echo=FALSE}
question("Which of the following statements apply to the figure above? (select all that apply)",
  answer("The larger the mole fraction of the chemical species the larger the chemical potential", correct = TRUE),
  answer("The higher the temperature the greater the chemical potential"),
  answer("Adding solutes reduces the mole fraction of our chemical species, which decreases the activity of our species, causing a reduction in chemical potential", correct = TRUE),
  allow_retry = TRUE
)
```

Create a bar plot that compares the chemical potential of water in a root and in a branch of a tree. Show the chemical potential of each term as well as their totals. Assume the following environmental conditions:

* Absolute temperature: 295 K
* Water mole fraction: 0.999 (-)
* Absolute minus atmospheric pressure in root: -0.1 x 10^5^ Pa  
* Absolute minus atmospheric pressure in stem: -3.0 x 10^5^ Pa
* Root depth: 0.5 m below ground
* Branch height: 10 m above ground
  
```{r chemPot, exercise = TRUE, exercise.eval = FALSE}
# constants
R <- ___ # universal gas constant (J K–1 mol–1)
g <- ___ # gravitational acceleration (m s-2)

# variables
T <- ___ # absolute temperature (K)
c <- ___ # mole fraction (-)
V <- ___ # partial molal volume (m3 mol-1)
m <- ___ # molar mass (kg mol-1)

# root
p <- ___ # absolute minus atmospheric pressure (Pa)
h <- ___ # height above ground (m)

activity <- ___
pressure <- ___
height <- ___
total <- ___
chemPot.root <- c(activity, pressure, height, total)

# branch
p <- ___ # absolute minus atmospheric pressure (Pa)
h <- ___ # height above ground (m)

activity <- ___
pressure <- ___
height <- ___
total <- ___
chemPot.branch <- c(activity, pressure, height, total)

data <- cbind(chemPot.root, chemPot.branch)

rownames(data) <- c("activity", "pressure", "height", "total")
colnames(data) <- c("root","branch")

my.ylab <- "Chemical potential of water (J/mol)"

barplot(data, ylim = c(-8, 5), ylab = my.ylab, beside=TRUE, legend = TRUE, args.legend = c(x="top", horiz = TRUE, bty = 'n'))

```

```{r chemPot-solution}
# constants
R <- 8.314 # universal gas constant (J K–1 mol–1)
g <- 9.82 # gravitational acceleration (m s-2)

# variables
T <- 295 # absolute temperature (K)
c <- 0.999 # mole fraction (-)
V <- 18 * 10^{-6} # partial molal volume (m3 mol-1)
m <- 10 * 10^{-3} # molar mass (kg mol-1)

# root
p <- -0.1 * 100000 # absolute minus atmospheric pressure (Pa)
h <- -0.5 # height above ground (m)

activity <- R * T * log(c)
pressure <- V * p
height <- m * g * h
total <- activity + pressure + height
chemPot.root <- c(activity, pressure, height, total)

# branch
p <- -3 * 100000 # absolute minus atmospheric pressure (Pa)
h <- 10 # height above ground (m)

activity <- R * T * log(c)
pressure <- V * p
height <- m * g * h
total <- activity + pressure + height
chemPot.branch <- c(activity, pressure, height, total)

data <- cbind(chemPot.root, chemPot.branch)

rownames(data) <- c("activity", "pressure", "height", "total")
colnames(data) <- c("root","branch")

my.ylab <- "Chemical potential of water (J/mol)"

barplot(data, ylim = c(-8, 5), ylab = my.ylab, beside=TRUE, legend = TRUE, args.legend = c(x="top", horiz = TRUE, bty = 'n'))

```

```{r chemPotQuestions, echo=FALSE}
question("Which of the following statements apply to the figure above? (select all that apply)",
  answer("The water flows from the root to the branch because of differences in the pressure term", correct = TRUE),
  answer("The water cannot flow from the root to the branch because of differences in the gravitational term"),
  answer("The gravitational term of the water in the root is negative", correct = TRUE),
  answer("The activity term in roots and branches are equal if the solute concentration is constant", correct = TRUE),
  allow_retry = TRUE
)
``` 

### Water potential
The water potential can be defined as: $$ \Psi_{w}=\Psi_{\pi}+\Psi_{p}+\Psi_{g}+\Psi_{m}$$
```{r sortWaterPotentialTerms, echo = FALSE}

terms <- c("Osmotic potential", "Pressure potential", "Gravity potential", "Matric potential")
rank_terms <- sortable::question_rank(
  "Sort the terms in the correct order given by the equation above",
  learnr::answer(terms, correct = TRUE),
  learnr::answer(rev(terms), correct = FALSE, "Other direction!"),
    allow_retry = TRUE
)
rank_terms
```

```{r waterPotQuestions, echo=FALSE}
question("Which of the following statements is correct? (select all that apply)",
answer("Osmotic potential is less or equal to zero", correct = TRUE),
answer("Adding solutes reduces the osmotic potential", correct = TRUE),
answer("Pressure potential is always negative"),
answer("Tension reduces the pressure potential", correct = TRUE),
answer("Gravity potential is always positive"),
answer("Lifting water increases gravity potential", correct = TRUE),
answer("Matric potential is less or equal to zero", correct = TRUE),
answer("Soil matrices reduce matric potential", correct = TRUE),
  allow_retry = TRUE
)
```
