
## Chemical Reactions
### Reaction kinetics
Consider the following reaction: $$v=k[A][B][C]^2$$
```{r reactionOrder, echo=FALSE}
question_numeric("What is the corresponding reaction order?",
answer("4", correct = TRUE),
  allow_retry = TRUE
)
```

```{r reactionRate, echo=FALSE}
question_text("Which variable represents the reaction rate?",
answer("v", correct = TRUE),
  allow_retry = TRUE
)
```

```{r reactionConstantUnit, echo=FALSE}
question_radio("What is the unit of the reaction rate constant?",
answer("mol<sup>3</sup> l<sup>-3</sup> s<sup>-1</sup>"),
answer("mol<sup>3</sup> l<sup>-3</sup> s<sup>1</sup>"),
answer("mol<sup>-3</sup> l<sup>3</sup> s<sup>-1</sup>", correct = TRUE),
answer("mol<sup>-4</sup> l<sup>-4</sup> s<sup>-1</sup>"),
  allow_retry = TRUE
)
```

```{r ChemicalEquilibrium, echo=FALSE}
question("Which of the following statements on chemical equilibrium are true? (select all that apply)",

answer("The concentrations of reactants and products are equal"),

answer("The concentrations of reactants and products are constant because forward and backward reactions stop"),

answer("The concentrations of reactants and products are constant because forward and backward reaction rates are equal", correct = TRUE),

answer("Reversible reactions continue until all reactants convert to products"),

answer("Reversible reactions always tend to come to equilibrium", correct = TRUE),

answer("When the system is at equilibrium, the amount of products and the reactants have not to be necessarily equal", correct = TRUE),

answer("The reaction rate konstants of the forward and backward reactions are equal"),

answer("The reaction rates of the forward and backward reactions are equal", correct = TRUE),

allow_retry = TRUE
)
```

```{r equilibriumConstant, echo=FALSE}
question("What are the implications if the equilibrium constant is greater than one? (select all that apply)",

answer("The products are in greater concentration than the reactants at equilibrium and the reaction tends to go to completion", correct = TRUE),

answer("The reactants are in greater concentration than the products at equilibrium and the reaction tends to go to completion"),

answer("The equilibrium constant never exceeds a value of one"),

allow_retry = TRUE
)
```

Consider the reaction: $$ CO_2+H_2\underset{\text{k$_2$}}{\overset{\text{k$_1$}}{\rightleftarrows}}CO+H_2O $$
Calculate the value of the equilibrium constant for the system shown for 0.0954 moles/l of CO$_2$, 0.0454 moles/l of H$_2$, 0.0046 moles/l of CO, and 0.0046 moles/l of H$_2$O.

```{r equilibriumConstantExample, exercise = TRUE, exercise.eval = FALSE}

```

```{r equilibriumConstantExample-solution}
CO2 <- 0.0954
H2 <- 0.0454
CO <- 0.0046
H2O <- 0.0046
Ke <- CO2 * H2 / (CO / H2)
print(Ke)

```

```{r equilibriumConstantExample-check}
grade_result(
  pass_if(~identical(round(.result, 4), 0.0427))
)
```

### The energetics of chemical reactions

Show how the fraction of nitrogen gas molecules of a given speed changes when the temperature increases from 300 to 500 K. Recall that the Maxwell-Boltzmann energy distribution is given as:

$$ F_c(c) = 4 \pi c^2 \left( \frac{m}{2 \pi k_B T} \right)^{\frac{3}{2}} \exp{\frac{-m c^2}{2 k_B T}} $$ Where $c$ is the speed of a molecule (m s$^{-1}$), $m$ is the mass of a single molecule (kg) and $k_B$ is the Boltzmann constant ($k_B = 1.38 \times 10^{-23}$ J K$^{-1}$)

```{r MaxwellBoltzmann, exercise = TRUE, exercise.eval = FALSE}
# constants
Avogadro <- 6.02 * 10^(23) # particles per mol 
kB <- 1.38 * 10^(-23) # Boltzmann constant (J/K)

# inputs
molarMass <- ___ # molar mass of N2 in g/mol
T <- ___ # temperature in K
c <- ___ # speed in m/s
m <- ___ # mass of a N2 molecule in kg

my.fun <- function(c, T){

# Fraction of molecules with a given speed  
y <- ___
  return(y)
}

data <- outer(c, T, my.fun)

my.xlab <- "Molecular speed $c$ in m s$^{-1}$"
my.ylab <- "Fraction of molecules (-)"
my.xlab <- latex2exp::TeX(my.xlab)

my.col <- viridis::plasma(n = ncol(data))

matplot(x = ___, y = ___, type = "l", lty = 1, col = my.col,
        xlab = my.xlab, ylab = my.ylab)

legend.text <- paste("T =", T, "K", sep = " ")
legend("topright", legend = legend.text, pch = 16, col = my.col, bty = "n")
```

```{r MaxwellBoltzmann-solution}
# constants
Avogadro <- 6.02 * 10^(23) # particles per mol 
kB <- 1.38 * 10^(-23) # Boltzmann constant (J/K)

# inputs
molarMass <- 28.014 # molar mass of N2 in g/mol
T <- seq(300, 500, 100) # temperature in K
c <- seq(1, 1501, 10) # speed in m/s
m <- molarMass / Avogadro / 1000 # mass of a N2 molecule in kg

my.fun <- function(c, T){
# Fraction of molecules with a given speed  
y <- 4 * pi * c^2 * (m / (2 * pi * kB * T))^(3/2) * exp( -1 * m * c^2 / (2 * kB * T))
  return(y)
}

data <- outer(c, T, my.fun)

my.xlab <- "Molecular speed $c$ in m s$^{-1}$"
my.ylab <- "Fraction of molecules (-)"
my.xlab <- latex2exp::TeX(my.xlab)

my.col <- viridis::plasma(n = ncol(data))

matplot(x = c, y = data, type = "l", lty = 1, col = my.col,
        xlab = my.xlab, ylab = my.ylab)

legend.text <- paste("T =", T, "K", sep = " ")
legend("topright", legend = legend.text, pch = 16, col = my.col, bty = "n")

```


Make a similar Figure as the one above, except that you express the fraction of molecules at a given speed as a function of their kinetic energy.
```{r kineticEnergy, exercise = TRUE, exercise.eval = FALSE}

```

```{r kineticEnergy-solution}
# constants
Avogadro <- 6.02 * 10^(23) # particles per mol 
kB <- 1.38 * 10^(-23) # Boltzmann constant (J/K)

# inputs
molarMass <- 28.014 # molar mass of N2 in g/mol
T <- seq(300, 500, 100) # temperature in K
c <- seq(1, 1501, 10) # speed in m/s
m <- molarMass / Avogadro / 1000 # mass of a N2 molecule in kg
kineticEnergy <- 0.5 * m * c^2 # kinetic energy

my.fun <- function(c, T){

# Fraction of molecules with a given speed  
y <- 4 * pi * c^2 * (m / (2 * pi * kB * T))^(3/2) * exp( -1 * m * c^2 / (2 * kB * T))
  return(y)
}

data <- outer(c, T, my.fun)

my.xlab <- "Kinetic Energy $E$ (J)"
my.ylab <- "Fraction of molecules (-)"
my.xlab <- latex2exp::TeX(my.xlab)

my.col <- viridis::plasma(n = ncol(data))

matplot(x = kineticEnergy, y = data, type = "l", lty = 1, col = my.col,
        xlab = my.xlab, ylab = my.ylab)

legend.text <- paste("T =", T, "K", sep = " ")
legend("topright", legend = legend.text, pch = 16, col = my.col, bty = "n")

```
For the figure above, will the majority of nitrogen gas molecules exceed an activation energy of $3 \times 10^{-20}$?

```{r activationEnergy, echo=FALSE}
question_radio(
  "According to the figure above, will the majority of nitrogen gas molecules exceed an activation energy of 3 x 10^(-20) J?",
  answer("Yes"),
  answer("No", correct = TRUE),
  allow_retry = TRUE
)
```

Make a Figure that shows how the rate constant $k$ varies with temperature for different values of activation energy. Allow temperatures to vary between 270K and 320K and activation energy to range from 50 to 55 kJ mol^-1^. Assume a frequency factor of $C = 10^{10}$ mol$^{-1}$ l s$^{-1}$.

```{r Arrhenius, exercise = TRUE, exercise.eval = FALSE}
# constants
R <- 8.314 # universal gas constant in J mol-1 K-1 

# variables
Ea <- ___ # activation energy in kJ/mol
T <- ___ # temperature in K
C <- ___ # frequency factor

my.fun <- function(T, Ea){
y <- ___ # Arrhenius equation
return(y)
}

data <- outer(___)

my.xlab <- "Temperature (K)"
my.ylab <- "Reaction rate constant k (mol$^{-1}$ l s$^{-1}$)"
my.ylab <- latex2exp::TeX(my.ylab)

my.col <- viridis::plasma(n = ncol(data))

matplot(x = T, y = data, type = "l", lty = 1, col = my.col,
        xlab = my.xlab, ylab = my.ylab)

legend.text <- paste("Ea =", Ea, "kJ/mol", sep = " ")
legend("topleft", legend = legend.text, pch = 16, col = my.col, bty = "n")
```

```{r Arrhenius-solution}
# constants
R <- 8.314 # universal gas constant in J mol-1 K-1 

# variables
Ea <- seq(50, 55, 1) # activation energy in kJ/mol
T <- seq(270, 320, 1) # temperature in K
C <- 10^10 # frequency factor

my.fun <- function(T, Ea){
y <- C * exp(-1 * Ea * 1000 / (R * T)) # Arrhenius equation
return(y)
}

data <- outer(T, Ea, my.fun)

my.xlab <- "Temperature (K)"
my.ylab <- "Reaction rate constant k (mol$^{-1}$ l s$^{-1}$)"
my.ylab <- latex2exp::TeX(my.ylab)

my.col <- viridis::plasma(n = ncol(data))

matplot(x = T, y = data, type = "l", lty = 1, col = my.col,
        xlab = my.xlab, ylab = my.ylab)

legend.text <- paste("Ea =", Ea, "kJ/mol", sep = " ")
legend("topleft", legend = legend.text, pch = 16, col = my.col, bty = "n")

```

```{r ArrheniusQuestion, echo=FALSE}
question_radio("How does the reaction rate constant change with temperature and activation energy?",
answer("The reaction rate constant increases with increasing temperature and increasing activation energy"),
answer("The reaction rate constant increases with increasing temperature and decreasing activation energy", correct = TRUE),
answer("The reaction rate constant is constant and does therefore not depend on temperature or activation energy"),
  allow_retry = TRUE
)
```

The $Q_{10}$ model is defined as: $$Q_{10}=\left(\frac{k_{T2}}{k_{T1}}\right)^{\frac{10}{T_2-T_1}}$$
Make a Figure that shows how the ratio of $k_{T2}$ and $k_{T1}$ changes with temperatures ranging from 270K to 320K. At what temperature does the ratio equal 2?
```{r Q10model, exercise = TRUE, exercise.eval = FALSE}
# constants 
R <- 8.314 # universal gas constant in J mol-1 K-1 

# variables
Ea <- ___ # activation energy in kJ/mol
T1 <- ___ # temperature in K
T2 <- ___ # temperature in K
C <- ___

k1 <- ___
k2 <-___

my.xlab <-  latex2exp::TeX("Temperature T$_1$ (K)")
my.ylab <- latex2exp::TeX("$k_2 / k_1$")

plot(___, ___, type = "l", xlab = my.xlab, ylab = my.ylab,
     tcl = 0.3)

abline(h = 2, lty = 3)
abline(v = 290, lty = 3)

```

```{r Q10model-solution}
# constants 
R <- 8.314 # universal gas constant in J mol-1 K-1 

# variables
Ea <- 50 # activation energy in kJ/mol
T1 <- seq(270, 320, 1) # temperature in K
T2 <- T1 + 10

C <- 1

k1 <- C * exp(-1 * Ea * 1000 / (R * T1))
k2 <- C * exp(-1 * Ea * 1000 / (R * T2))

my.xlab <-  latex2exp::TeX("Temperature T$_1$ (K)")
my.ylab <- latex2exp::TeX("$k_2 / k_1$")

plot(T1, k2/k1, type = "l", xlab = my.xlab, ylab = my.ylab,
     tcl = 0.3)

abline(h = 2, lty = 3)
abline(v = 290, lty = 3)

```

### Redox reaction

Consider the following redox reaction: $$ SO_3^{2-} + MnO_4^- \rightarrow SO_4^{2-}+Mn^{2+}$$

```{r redoxReaction_Mn, echo=FALSE}
question_numeric("By how much does the oxidation number of Manganese change?",
answer("-5", correct = TRUE),
  allow_retry = TRUE
)
```

```{r redoxReaction_S, echo=FALSE}
question_numeric("By how much does the oxidation number of Sulfure change?",
answer("2", correct = TRUE),
  allow_retry = TRUE
)
```

```{r redoxReaction, echo=FALSE}
question("Which of the following statements apply to the figure above? (select all that apply)",
  answer("Manganese is being reduced", correct = TRUE),
  answer("Manganese is being oxidized"),
  
  answer("Sulfure is being reduced"),
  answer("Sulfure is being oxidizd", correct = TRUE),
  
  answer("Manganese accepts electrons", correct = TRUE),
  answer("Manganese donates electrons"),
  
  answer("Sulfure donates electrons", correct = TRUE),
  answer("Sulfure accepts electrons"),

  allow_retry = TRUE
)
``` 

Consider the following redox couples and their corresponding standard reduction potentials ($E^o$):
$$Cu^{2+} + 2e^- \rightarrow Cu; E^o = +0.340V$$
$$Zn^{2+} + 2e^- \rightarrow Zn; E^o = -0.763V$$
```{r redoxPotentials, echo=FALSE}
question_radio("Which of these two redox couples is a better electron donor?",
answer("Cu2+"),
answer("Zn2+", correct = TRUE),
  allow_retry = TRUE
)
```

### Enzyme catalysis

Create a Figure that shows how the reaction rate varies with substrate concetration (ranging from 0 to 5000 $\mu$mol/l) for various Michaelis constants (ranging from 500 to 1000). Assume a maximum reaction rate of 50 x 10$^{-6}$ mol l$^{-1}$ s$^{-1}$. 

```{r MichaelisMenten, exercise = TRUE, exercise.eval = FALSE}

# inputs
Vmax <- ___ # maximum reaction rate (x 10^-6 mol l-1 s-1)
Km <- ___ # equillibrium constant
S <- ___ # substrate in micro mol/l

my.fun <- function(S, Km){
  y <- ___ # reaction rate
  return(y)
}

  myColName <- paste("K_m = ",Km," (10$^{-6}$ mol l$^{-1}$)", sep = "")
  
data <- outer(S, Km, my.fun)

my.xlab <- "Substrate concentration S (10$^{-6}$ mol l$^{-1}$)"
my.ylab <- "Reaction rate v (10$^{-6}$ mol l$^{-1}$ s$^{-1}$)"

my.xlab <- latex2exp::TeX(my.xlab)
my.ylab <- latex2exp::TeX(my.ylab)

my.col <- viridis::plasma(n = ncol(data))

matplot(x = S, y = data, type = "l", lty = 1, col = my.col,
        xlab = my.xlab, ylab = my.ylab)

legend.text <- paste("K$_m$ = ",Km," (10$^{-6}$ mol l$^{-1}$)", sep = "")
legend.text <- latex2exp::TeX(legend.text)

legend("bottomright", legend = legend.text, pch = 16, col = my.col, bty = "n")

```

```{r MichaelisMenten-solution}

# inputs
Vmax <- 50 # maximum reaction rate (x 10^-6 mol l-1 s-1)
Km <- seq(500, 1000, 250)
S <- seq(0, 5000, 10) # substrate in micro mol/l

my.fun <- function(S, Km){
  y <- Vmax * S / (Km + S) # reaction rate
  return(y)
}

  myColName <- paste("K_m = ",Km," (10$^{-6}$ mol l$^{-1}$)", sep = "")
  
data <- outer(S, Km, my.fun)

my.xlab <- "Substrate concentration S (10$^{-6}$ mol l$^{-1}$)"
my.ylab <- "Reaction rate v (10$^{-6}$ mol l$^{-1}$ s$^{-1}$)"

my.xlab <- latex2exp::TeX(my.xlab)
my.ylab <- latex2exp::TeX(my.ylab)

my.col <- viridis::plasma(n = ncol(data))

matplot(x = S, y = data, type = "l", lty = 1, col = my.col,
        xlab = my.xlab, ylab = my.ylab)

legend.text <- paste("K$_m$ = ",Km," (10$^{-6}$ mol l$^{-1}$)", sep = "")
legend.text <- latex2exp::TeX(legend.text)

legend("bottomright", legend = legend.text, pch = 16, col = my.col, bty = "n")

```

```{r MichaelisMenten_firstOrderKinetics, echo=FALSE}
question_radio("When is the reaction rate approximately first-order kintics?",
answer("When [S] >> Km"),
answer("When [S] << Km", correct = TRUE),
answer("When [S] = Km"),
  allow_retry = TRUE
)
```

```{r MichaelisMenten_zeroOrderKinetics, echo=FALSE}
question_radio("When is the reaction rate approximately zero-order kintics?",
answer("When [S] >> Km", correct = TRUE),
answer("When [S] << Km"),
answer("When [S] = Km"),
  allow_retry = TRUE
)
```

```{r MichaelisMenten_Km, echo=FALSE}
question("Which of the following statements are correct? (select all that apply)",
  answer("The lesser the value of Km, the tighter the association between enzyme and substrate", correct = TRUE),
  answer("The larger the value of Km, the tighter the association between enzyme and substrate"),
  answer("Km is the substrate concentration that yields half Vmax", correct = TRUE),
  allow_retry = TRUE
)
``` 