## Water transport
### Water transport through soils

```{r matricPotentialSign, echo=FALSE}
question_radio(
  "The matric potential of unsaturated soils is:",
  answer("positive"),
  answer("negative", correct = TRUE),
  allow_retry = TRUE
)
```

```{r retention01, echo=FALSE}
question_radio(
  "An increase in volumetric soil water content causes matric potential to become",
  answer("more negative"),
  answer("less negative", correct = TRUE),
  allow_retry = TRUE
)
```

```{r retention02, echo=FALSE}
question_radio(
  "An increase in grain size causes matric potential to become",
  answer("more negative"),
  answer("less negative", correct = TRUE),
  allow_retry = TRUE
)
```

```{r conductivity01, echo=FALSE}
question_radio(
  "An increase in volumetric soil water content causes conductivity to",
  answer("increase", correct = TRUE),
  answer("decrease"),
  allow_retry = TRUE
)
```

```{r conductivity02, echo=FALSE}
question_radio(
  "An increase in grain size causes conductivity to",
  answer("increase", correct = TRUE),
  answer("decrease"),
  allow_retry = TRUE
)
```

```{r retentionAndConductivity, echo=FALSE}
question_radio(
  "The retention curve shows the relation between volumetric soil water content and",
  answer("Matric potential", correct = TRUE),
  answer("Soil water potential"),
  answer("Hydraulic conductivity"),
  allow_retry = TRUE
)
```

Consider a soil with a water potential of -0.05 MPa. What is the corresponding water potential expressed in J mol^-1^ assuming a temperature of 25^o^C and an atmospheric pressure of 1000 hPa?

```{r PsiUnitConversion01, exercise = TRUE, exercise.eval = FALSE}

```

```{r PsiUnitConversion01-solution}
psi_Pa <- -0.05 * 10^6 # water potential in Pa
partialMolalVolume <- 18 * 10^(-6) # partial molal volume (m3 mol-1)
psi_JPerMol <- psi_Pa * partialMolalVolume # water potential in J/mol
print(psi_JPerMol)
```

```{r PsiUnitConversion01-check}
grade_result(
  pass_if(~identical(round(.result), round(-0.9)))
)

```

Repeat the same exercise as above but converting the units from J mol^-1^ to m.

```{r PsiUnitConversion02, exercise = TRUE, exercise.eval = FALSE}

```

```{r PsiUnitConversion02-solution}
psi_Pa <- -0.05 * 10^6 # water potential in Pa
rho <- 1000 # water density (kg m-3)
g <- 9.81 # gravitational acceleration (m s-2)

psi_m <- psi_Pa / (rho * g) # water potential in m
print(psi_m)
```

```{r PsiUnitConversion02-check}
grade_result(
  pass_if(~identical(round(.result), round(-5.09684)))
)

```

The retention curve function by van Genuchten (1980) is

$$ \theta = \frac{\theta_{max} - \theta_{min}}{(1 + (a |\psi_m|)^n)^m} + \theta_{min}$$
$$ m = 1- 1/n $$

Plot a retention curve for a matric potential for sand and loam using the parameters below. Let the matric potential range from -1000 to -1 MPa.


| Parameter       | Sand    | Loam    | Units     |
|:----------------|:--------|:--------|:----------|
| a               | 0.028   |0.0104   |cm^-1^     |
| n               | 2.239   |1.3954   |(-)        |
| $\theta_{min}$  | 0.0286  |0.1060   |m^3^ m^-3^ |
| $\theta_{max}$  | 0.3658  |0.4686   |m^3^ m^-3^ |

```{r retentionCurve, fig.width = 4, fig.height = 6, exercise = TRUE, exercise.eval = FALSE}

# To use a logarithmic axis use log = "y" in your plotting function, e.g.
x <- seq(1,100,1)
y <- x^2
plot(x, y, log = "y", type = "l")

```

```{r retentionCurve-solution, fig.width = 4, fig.height = 6}
rm(list=ls())

# variable
psi_m <- seq(1,1000,1) # positive values

# sand:
a <- 0.028 # parameter (cm-1)
n <- 2.239 # parameter (-)
theta_min <- 0.0286
theta_max <- 0.3658

m <- 1 - 1/n
denom <- (1 + (a * abs(psi_m))^n)^m
theta <- (theta_max - theta_min) * 1/denom + theta_min
theta.sand <- theta

# loam:
a <- 0.0104 # parameter (cm-1)
n <- 1.3954 # parameter (-)
theta_min <- 0.106
theta_max <- 0.4686

m <- 1 - 1/n
denom <- (1 + (a * abs(psi_m))^n)^m
theta <- (theta_max - theta_min) * 1/denom + theta_min
theta.loam <- theta

my.xlab <- "Volumetric soil water content $\\theta_s$ (m$^3$ m$^{-3}$)"
my.ylab <- "Negative Matric potential $\\psi$ (MPa)"
my.xlab <- latex2exp::TeX(my.xlab)
my.ylab <- latex2exp::TeX(my.ylab)

plot(x = theta.sand, y = psi_m, xlim = c(0, 0.6), log = "y", type = "l", xlab = my.xlab, ylab = my.ylab)
lines(x = theta.loam, y = psi_m, lty = 2)
legend("topright", lty = c(1,2), c("sand", "loam"), bty = 'n')

```

The hydraulic conductivity function by van Genuchten (1980) is

$$ K = K_{sat} \times S_e^{1/2}\left( 1 - \left( 1 - S_e^{1/m} \right)^m \right)^2$$

$$ S_e = \frac{1}{(1 + (a |\psi_m|)^n)^m}$$
$$ m = 1- 1/n $$

Plot the hydraulic conductivity for sand and loam using the parameters below. Let the matric potential range from -1000 to -1 MPa.

| Parameter       | Sand    | Loam    | Units     |
|:----------------|:--------|:--------|:----------|
| a               | 0.028   |0.0104   |cm^-1^     |
| n               | 2.239   |1.3954   |(-)        |
| $\theta_{min}$  | 0.0286  |0.1060   |m^3^ m^-3^ |
| $\theta_{max}$  | 0.3658  |0.4686   |m^3^ m^-3^ |
| $K_{sat}$       | 29.7    |1.04     |cm h^-1^   |

```{r hydraulicConductivity, fig.width = 4, fig.height = 6, exercise = TRUE, exercise.eval = FALSE}

```

```{r hydraulicConductivity-solution, fig.width = 4, fig.height = 6}
rm(list=ls())

# variable
psi_m <- seq(1,1000,1) # positive values

# sand:
a <- 0.028 # parameter (cm-1)
n <- 2.239 # parameter (-)
theta_min <- 0.0286
theta_max <- 0.3658
K_sat <- 29.7 # parameter (cm h-1)

m <- 1 - 1/n
denom <- (1 + (a * abs(psi_m))^n)^m
theta <- (theta_max - theta_min) * 1/denom + theta_min
theta.sand <- theta

Se <- 1 / denom
K <- K_sat * Se^(0.5) * (1 - (1 - Se^(1/m))^m)^2
K.sand <- K

# loam:
a <- 0.0104 # parameter (cm-1)
n <- 1.3954 # parameter (-)
theta_min <- 0.106
theta_max <- 0.4686
K_sat <- 1.04 # parameter (cm h-1)

m <- 1 - 1/n
denom <- (1 + (a * abs(psi_m))^n)^m
theta <- (theta_max - theta_min) * 1/denom + theta_min
theta.loam <- theta
Se <- 1 / denom
K <- K_sat * Se^(0.5) * (1 - (1 - Se^(1/m))^m)^2
K.loam <- K

my.xlab <- "Volumetric soil water content $\\theta_s$ (m$^3$ m$^{-3}$)"
my.ylab <- "Hydraulic conductivity K (cm h$^{-1}$)"
my.xlab <- latex2exp::TeX(my.xlab)
my.ylab <- latex2exp::TeX(my.ylab)

plot(x = theta.sand, y = K.sand, xlim = c(0, 0.6), log = "y", type = "l", xlab = my.xlab, ylab = my.ylab)
lines(x = theta.loam, y = K.loam, lty = 2)
legend("bottomright", lty = c(1,2), c("sand", "loam"), bty = 'n')

```

### Water transport through roots

```{r sortAnatomicalStructureOfRoots, echo = FALSE}

units <- c("epidermis", "cortex", "endodermis", "vascular tissue")
rank_units <- sortable::question_rank(
  "Sort the parts of the roots as they appear going from the outside towards the center of the root",
  learnr::answer(units, correct = TRUE),
  learnr::answer(rev(units), correct = FALSE, "Other direction!"),
  allow_retry = TRUE
)
rank_units
```

```{r rootsStorage, echo=FALSE}
question_radio(
  "Which part of the root stores organic polymers that can be used for future growth?",
  answer("epidermis"),
  answer("endodermis"),
  answer("vascular tissue"),
  answer("cortex", correct = TRUE),
  allow_retry = TRUE
)
```

```{r stele, echo=FALSE}
question("Which parts are encolsed by the stele? (select all that apply)",

answer("epidermis"),
answer("endodermis", correct = TRUE),
answer("cortex"),
answer("phloem", correct = TRUE),
answer("xylem", correct = TRUE),

allow_retry = TRUE
)
```

```{r flowDirection, echo=FALSE}
question_radio(
  "In which part of the root does the waer follow a radial path?",
  answer("xylem"),
  answer("phloem"),
  answer("cortex", correct = TRUE),
  allow_retry = TRUE
)
```

```{r apoplasticSymplastic, echo=FALSE}
question_radio(
  "Which of the two types of water flow through roots is driven by a gradient is osmotic potential?",
  answer("apoplastic flow"),
  answer("symplastic flow", correct = TRUE),
  allow_retry = TRUE
)
```

```{r casparianStrip, echo=FALSE}
question_radio(
  "Which of the two types of water flow through roots is blocked by the Casparian strip?",
  answer("apoplastic flow", correct = TRUE),
  answer("symplastic flow"),
  allow_retry = TRUE
)
```

### Water transport through stems

The decline of hydraulic conductance in relation to water potential for stems can be modeled based on a two-parameter Weibull function:

$$ K = K_{max} \exp \left( - \left( \frac{\psi}{d} \right)^c \right)$$

Where $K_{max}$ is the maximum hydraulic conductivity and $c$ and $d$ are empirical parameters. For plant (1) the parameter values are $c$ = 3.43 (-) and $d$ = -2.15 MPa. For plant (2) the parameter values are $c$ = 2.64 (-) and $d$ = -3.54 MPa. Make a figure that shows how the ratio ratio $K/K_{max}$ varies for both plants with xylem water potential ranging from -6 to 0 MPa. Discuss with your neighbor which of the two plants you suspect to the a conifer, and which to be a deciduous tree. 

```{r stemConductivity, exercise = TRUE, exercise.eval = FALSE}

```

```{r stemConductivity-solution}
rm(list=ls())

# variable
psi <- seq(-6, 0, 0.1 ) # xylem water potential 

# plant (1)
c <- 3.43 # empirical parameter (-)
d <- -2.15 # empirical parameter (MPa)
ratio <- exp(- (psi/d)^c) # K/K_max
plant01 <- ratio

c <- 2.64 # empirical parameter (-)
d <- -3.54 # empirical parameter (MPa)
ratio <- exp(- (psi/d)^c) # K/K_max
plant02 <- ratio

my.xlab <- "Stem water potential $\\psi$ (MPa)"
my.ylab <- "$K/K_{max}$ (-)"
my.xlab <- latex2exp::TeX(my.xlab)
my.ylab <- latex2exp::TeX(my.ylab)

plot(x = psi, y = plant01, type = "l", xlab = my.xlab, ylab = my.ylab)
lines(x = psi, y = plant02, lty = 2)
legend("topleft", c("Deciduous", "Conifer"), lty = c(1, 2), bty = "n")

```

Discuss with your neighbor why and in what circumstances Darcy's law may overestimate the stem water flow observed in plants.
