## Energy budgets

### Net radiation

Calculate the net radiation of a leaf given the following inputs:

| Variable      | Value     | Units              |
|:--------------|:----------|:-------------------|
| $R_s$         | 161       |W m^-2^             |
| $R_{Lu}$      | 333       |W m^-2^             |
| $R_{Ll}$      | 396       |W m^-2^             |
| $a_{su}$      | 0.8       |(-)                 |
| $a_{sl}$      | 0.8       |(-)                 |
| $\alpha_g$    | 0.3       |(-)                 |
| $\varepsilon$ | 0.9       |(-)                 |
| $T_s$         | 20        |(^o^C)              |

```{r netRadiation01, exercise = TRUE}

```

```{r netRadiation01-solution}
# constants:
epsilon <- 0.9
sigma <- 5.67 * 10^(-8) # Stefan-Boltzmann constant (W m-2 K-4)

# Variables:
Rs <- 161 # incoming SW radiation (W m-2)
R_Lu <- 333 # LW radiation coming from above (W m-2)
R_Ll <- 396 # LW radiation coming from below (W m-2)
a_su <- 0.8 # absorptance of SW radiation of upper leaf surface (-)
a_sl <- 0.8 # absorptance of SW radiation of lower leaf surface (-)
aL <- epsilon # absorptance of LW radiation
alpha_g <- 0.3 # albedo of ground surface (-)

Ts <- 20 + 273.15 # leaf surface temperature (K)

Rn <- Rs * a_su + Rs * a_sl * alpha_g + R_Lu * aL + R_Ll * aL - 2 * epsilon * sigma * Ts^4
print(Rn)
```

```{r netRadiation01-check}
grade_result(
    pass_if(~identical(round(.result), round(69.81112)))
)
```

Using the same inputs as above, show how the leaf net radiation changes as leaf temperature increases from 10 to 25^o^C. Explain your result.


```{r netRadiation02, exercise = TRUE}

```

```{r netRadiation02-solution}
# constants:
epsilon <- 0.9
sigma <- 5.67 * 10^(-8) # Stefan-Boltzmann constant (W m-2 K-4)

# Variables:
Rs <- 161 # incoming SW radiation (W m-2)
R_Lu <- 333 # LW radiation coming from above (W m-2)
R_Ll <- 396 # LW radiation coming from below (W m-2)
a_su <- 0.8 # absorptance of SW radiation of upper leaf surface (-)
a_sl <- 0.8 # absorptance of SW radiation of lower leaf surface (-)
aL <- epsilon # absorptance of LW radiation
alpha_g <- 0.3 # albedo of ground surface (-)

Ts <- seq(10, 25, 1) + 273.15 # leaf surface temperature (K)

Rn <- Rs * a_su + Rs * a_sl * alpha_g + R_Lu * aL + R_Ll * aL - 2 * epsilon * sigma * Ts^4

my.xlab <- "Leaf surface temperature (K)"
my.ylab <- "Leaf net radiation $R_n$ (W m$^{-2}$)"
my.xlab <- latex2exp::TeX(my.xlab)
my.ylab <- latex2exp::TeX(my.ylab)

par(mar = c(5,5,1,1))
plot(x = Ts, y = Rn, type = "l", xlab = my.xlab, ylab = my.ylab)

```

### Turbulent heat fluxes

Consider a canopy with a latent heat flux of 50 W m^-2^. What is the corresponding evapotranspiration expressed in mm per day for an air temperature of 23^o^C? 

```{r latentHeatOfVap01, exercise = TRUE}

```

```{r latentHeatOfVap01-solution}
Ta <- 23
latentHeat <- 50 # latent heat flux (W m-2)

lambdaW <- 2.502 * 10^6 - (2.308 * 10^3 * Ta)

ET <- latentHeat / lambdaW # evapotranspiration (kg m-2 s-1)
ET <- ET * 86400 # evapotranspiration (kg m-2 day-1)
print(ET)
```

```{r latentHeatOfVap01-check}
grade_result(
      pass_if(~identical(round(.result,2), round(1.764046,2)))
)
```

Repeat the same exercise as above but allow for temperature to range from 5 to 30^o^C.

```{r latentHeatOfVap02, exercise = TRUE}

```

```{r latentHeatOfVap02-solution}
Ta <- seq(5, 30, 1)
latentHeat <- 50 # latent heat flux (W m-2)

lambdaW <- 2.502 * 10^6 - (2.308 * 10^3 * Ta)

ET <- latentHeat / lambdaW # evapotranspiration (kg m-2 s-1)
ET <- ET * 86400 # evapotranspiration (kg m-2 day-1)

my.xlab <- "Air temperature ($^o$C)"
my.ylab <- "Evapotranspiration ET (mm day$^{-1}$)"
my.xlab <- latex2exp::TeX(my.xlab)
my.ylab <- latex2exp::TeX(my.ylab)

par(mar = c(5,5,1,1))
plot(x = Ta, y = ET, type = "l", xlab = my.xlab, ylab = my.ylab)

```

Plot the Clausius-Clapeyron relationship for a temperature range of 0 to 45^o^C using Equation 10.8 on p. 230. Discuss with your neighbor the implications of the Clausius-Clapeyron relationship for precipitation in response to global warming.


```{r ClausiusClapeyron, exercise = TRUE}

```

```{r ClausiusClapeyron-solution}
rm(list=ls())

# inputs
a <- 611 # Pa
b <- 17.502 # (-)
c <- 240.97 # C
P <- 101300 # Pa
T <- seq(0,45,1) # air temperature in C
es <- a * exp(b * T / (T + c)) # saturation vapor pressure 
c_aw.star <- es/P

# plot
my.ylab <- latex2exp::TeX("c$_{aw}$ (mol mol$^{-1}$)")
my.xlab <- latex2exp::TeX("Temperature (Celsius)")

par(mar = c(5,5,1,1))
plot(T, c_aw.star, type = "l", xlab = my.xlab, ylab = my.ylab, las = 0,
     tcl = 0.3)
axis(3, labels = FALSE, tcl = 0.3)
axis(4, labels = FALSE, tcl = 0.3)

```

Consider a sample of air with a temperature of 40^o^C and a water vapor mole fraction of 0.023 mol mol^-1^. Based on the Figure above, by how much does the air need to cool down to reach the corresponding dew point temperature?

```{r dewPointTemperature, echo=FALSE}
question_radio(
  "Consider a sample of air with a temperature of 40^o^C and a water vapor mole fraction of 0.023 mol/mol. Based on the Figure above, by how many degrees Celsius does the air need to cool down to reach the corresponding dew point temperature?",
  answer("10"),
  answer("20", correct = TRUE),
  answer("30"),
  answer("40"),
  allow_retry = TRUE
)
```

```{r relativeHumidity01, echo=FALSE}
question_numeric("How large does the relative humidity (%) need to be so that the wet bulb temperature equals the air temperature?",
answer("100", correct = TRUE),
  allow_retry = TRUE
)
```

```{r saturationDeficit, echo=FALSE}
question_numeric("What is the saturation vapor deficit expressed in mol/mol if the water vapor mole fraction equals 0.02 and the saturation vapor model fraction equals 0.06 mol/mol?",
answer("0.04", correct = TRUE),
  allow_retry = TRUE
)
```

```{r relativeHumidity02, echo=FALSE}
question_numeric("What is the corresponding relative humidity expressed in percentage?",
answer("50", correct = TRUE),
  allow_retry = TRUE
)
```

```{r saturation, echo=FALSE}
question("How can an air parcel achieve saturation? (Select all that apply)",
answer("Heating"),
answer("Cooling", correct = TRUE),
answer("Adding moisture", correct = TRUE),
answer("Extracting moisture"),
answer("Cooling and adding mositure", correct = TRUE),
  allow_retry = TRUE
)
```

```{r wetBulbTemperature, echo=FALSE}
question_radio(
  "What needs to happen so that the air temperature of an air parcel reduces to the wet bulb temperature?",
answer("Only cooling"),
answer("Only adding moisture"),
answer("Cooling and adding mositure", correct = TRUE),
  allow_retry = TRUE
)
```


### Penman-Monteith equation

Calculate the latent heat flux using the Penman-Monteith equation and data provided below. Which term is larger, the diabatic component or the adiabatic component?

$$\lambda_wE = \frac{s }{s + \gamma*}(R_n - H_G) + \frac{\rho_m c_p(c_{aw}^*(T_a) - c_{aw}(T_a))}{(s+ \gamma^*)r_h}$$

* $s$ is the slope of saturation vapor pressure curve at air temperature T (kPa ^o^C^-1^)
* $T_a$ is the air temperature (^o^C)
* $c_p$ is the molar specific heat of dry air (J mol$^{-1}$ K$^{-1}$) (29.1 J mol$^{-1}$ K$^{-1}$)
* $\rho_m$ is the molar density of dry air (mol m$^{-3}$): $\rho_m=P/(RT_a)$
* $c_{aw}$ is the mole fraction of water vapor in the sample of air (-)
* $c_{aw}^*$ is the saturated water vapor mole fraction in the same sample of air (-)
* $\gamma*= \gamma \frac{r_{tw}}{r_h}= 0.93 \gamma$ (K^-1^)
* $\gamma$ is the psychrometric constant in K$^{-1}$: $\gamma = c_p/\lambda_w$ (K^-1^)
* $\lambda_w$ is the latent heat of vaporization $\lambda_w$ (MJ kg$^{-1}$): $\lambda_w = 2.502 \times 10^6 - (2.308 \times 10^3 \times T_a)$
* $r_h$ is the resistance to sensible heat transfer between the air and surface (s m$^{-1}$): $r_h = 1/g_{bh}$
* $g_{bh}$ is the boundary layer conductance for heat: $g_{bh} = \frac{K_{dh}}{z_{bl}}$ (m s^-1^)
* $z_{bl}$ is the depth of the boundary layer: $z_{bl} = 4 \times 10^{-3}\sqrt{d/u}$ (m)
* $K_{dh}$ is the the diffusion coefficient for heat (m$^2$ s$^{-1}$): $K_{dh} = K_0 \frac{P_0}{P} \left(\frac{T}{T_0} \right)^{1.81}$
* $P_0$ is the standard pressure ($P_0$ = 101325 Pa)
* $K_0$ is $18.9 \times 10^{-6}$ m$^{2}$ s$^{-1}$ (m^-2^ s^-1^)
* $d$ is leaf length scale (m)



| Symbol| Meaning                                                 | Value                   | Units           |
|:------|:--------------------------------------------------------|:------------------------|:----------------|
|	R 	  |	 universal gas constant R 	                            |	8.31	                  |	J K^-1^ mol^-1^	|
|	T~0~  |	 reference temperature 	                                |	273.15	                |	K	              |
|	P~0~  |	 reference pressure 	                                  |	101325	                |	Pa	            |
|	K~0~ 	|	 thermal diffusivity at standard pressure for 273 K 	  |	18.9 x 10^-6^ 	        |	m2 s^-1^	      |
|	T 	  |	 air temperature 	                                      |	273.15 + 14	            |	K	              |
|	P 	  |	 surface pressure 	                                    |	100000	                |	Pa	            |
|	u 	  |	 wind speed 	                                          |	1	                      |	m s^-1^	        |
|	d 	  |	 characteristic leaf dimension 	                        |	0.02	                  |	m	              |
|	c~p~ 	|	 molar specific heat of dry air 	                      |	1005 x 28.97 x 10^-3^	  |	J mol^-1^ K^-1^	|
|	c~aw~	|	 water vapor mole fraction 	                            |	0.015	                  |	mol mol^-1^	    |
|	R~n~ 	|	 net surface radiation                                  |	162	                    |	W m^-2^	        |
|	H~G~ 	|	 soil heat flux 	                                      |	1	                      |	W m^-2^	        |

```{r PMeq, exercise = TRUE, exercise.eval = FALSE}

```

```{r PMeq-solution}
rm(list=ls())

# constants:
R <- 8.31 # universal gas constant R (J K-1 mol-1)
T0 <- 273.15 # Reference temperature (K)
P0 <- 101325 # Reference pressure (Pa)
K0 <- 18.9 * 10^(-6) # Thermal diffusivity at standard pressure for 273 K (m2 s-1)

# variables:
T <- 273.15 + 14 # air temperature (K)
P <- 100000 # surface pressure (Pa)
u <- 0.1 # wind speed (m s-1)
d <- 0.02 # characteristic leaf dimension (m)
cp <- 1005 * 28.97 * 10^-3 # specific heat of dry air in J K-1 kg-1 times Molar mass in kg mol-1 gives the molar specific heat of dry air (J mol–1 K–1)
c_aw <- 0.015 # water vapor mole fraction (mol mol-1)
Rn <- 162 # net surface radiation (W m-2)
HG <- 1 # soil heat flux (W m-2) 

# leaf boundary layer resistance for heat
K_dh <- K0 * P0 / P * (T / T0)^1.81 # diffusion coefficient for sensible heat (m2 s–1)
rho_m <- P / (R * T) # molar density of dry air (mol m^-3)
z_bl = 4 * 10^(-3) * (d/u)^0.5 # leaf boundary layer height (m) 
g_bh <- rho_m * K_dh / z_bl # boundary layer conductance for heat diffusion (mol m-2 s-1)
g_bh <- g_bh / rho_m # divide by molar density to change units from mol m-2 s-1 to m s-1
rh <- 1/g_bh # boundary layer resistance for heat (s m-1)

# saturation water vapor mole fraction (mol mol-1)
a <- 611 # Pa
b <- 17.502 # (-)
c <- 240.97 # C
TC <- T - 273.15 # air temperature in C
es <- a * exp(b * TC / (TC + c)) # saturation vapor pressure 
c_aw.star <- es/P

# slope of Clausius-Clapeyron equation
TC <- T - 273.15 - 0.5 # air temperature in C MINUS 0.5
es <- a * exp(b * TC / (TC + c)) # saturation vapor pressure 
c_aw.star01 <- es/P
TC <- T - 273.15 + 0.5 # air temperature in C PLUS 05
es <- a * exp(b * TC / (TC + c)) # saturation vapor pressure 
c_aw.star02 <- es/P
s <- c_aw.star02 - c_aw.star01

# gamma star
lambda_w <- 2.502 * 10^6 - (2.308 * 10^3 * TC) # latent heat of vaporization in J kg-1 (p. 224)
lambda_w <- lambda_w / 1000 # in J g-1
lambda_w <- lambda_w * 18.02 # in J mol-1 (molar mass of water: 18.02 g/mol H2O)
gamma <- cp / lambda_w 
gammaStar <- 0.93 * gamma # Montheith and Unsworth (2008), Eq. 13.20, p. 237

# Penman-Monteith equation

termI <- s/(s + gammaStar) * (Rn - HG)

numerator <- rho_m * cp * (c_aw.star - c_aw)
denominator <- (s + gammaStar) * rh

termII <- numerator / denominator

lambda_wE <- termI + termII

print(lambda_wE)

```

```{r PMeq-check}
grade_result(
  pass_if(~identical(round(.result), round(109.8903)))
)
```
