## Turbulent fluxes
### Static stability

```{r lapseRate01, echo=FALSE}
question_radio(
  "What is the sign of the lapse rate in the troposphere?",
  answer("positive", correct = TRUE),
  answer("negative"),
  allow_retry = TRUE
)
```

```{r lapseRate02, echo=FALSE}
question_radio(
  "What is the sign of the lapse rate in the stratosphere?",
  answer("positive"),
  answer("negative", correct = TRUE),
  allow_retry = TRUE
)
```

```{r lapseRate03, echo=FALSE}
question_radio(
  "Consider an atmosphere where the decrease in air temperature with height exceeds the decrease given by the dry adiabatic lapse rate. What is the static stability of the atmosphere?",
  answer("The atmosphere is statically stable"),
  answer("The atmosphere is statically unstable", correct = TRUE),
  answer("The atmosphere is statically neutral"),
  allow_retry = TRUE
)
```

```{r lapseRate04, echo=FALSE}
question_radio(
  "In a statically stable atmosphere, what happens to an air parcel when it is being pushed up?",
  answer("The parcel rises"),
  answer("The parcel sinks", correct = TRUE),
  answer("The parcel remains at its new location"),
  allow_retry = TRUE
)
```

Calculate the potential air temperature of a dry air parcel with a temperature of 290 K located 100 m above the surface.

```{r potentialAirTemp, exercise = TRUE, exercise.eval = FALSE}

```

```{r potentialAirTemp-solution}
# Eq. 12.1, p. 287) 

# inputs
z <- 100 # height in m
Ta <- 290

# Stable atmosphere for an unsaturated air parcel:
DALR <- 9.8 * 10^-3 # dry adiabatic lapse rate K/m
# Potential temperature
theta <- Ta + z * abs(DALR) # potential air temperature
print(theta)
```

```{r potentialAirTemp-check}
grade_result(
  pass_if(~identical(round(.result,1), round(290.98,1)))
)
```

```{r shear, echo=FALSE}
question_radio(
  "Where do you expect shear to be a more dominant source of turbulence in the Planetary boundary layer?",
  answer("In the surface layer", correct = TRUE),
  answer("Above the surface layer"),
  allow_retry = TRUE
)
```

At what height above the surface is the production of turbulence through shear roughly equal to that of buoyancy for the surface measurements provided below? 

* Friction velocity: 0.2 m s^-1^, 
* Turbulent heat flux: -0.05 K m s^-1^
* Virtual air temperature: 295 K?

```{r obukhovLength, exercise = TRUE, exercise.eval = FALSE}

```

```{r obukhovLength-solution}
# Eq. 12.11, p. 292) 

# inputs
uStar <- 0.2 # friction velocity (m s-1)
theta_vt <- 295 # virtual air temperature (K)
turbulentHeatFlux <- -0.05 # turbulent heat flux (K m s-1)

# constants
g <- 9.81 # gravitational acceleration (m s-2)
k <- 0.4 # van Karman constant (-)

# Calculate Obukhov length (m)

numerator <- uStar^3 * theta_vt
denominator <- g * k * turbulentHeatFlux
L <- (-1) * numerator / denominator
  
print(L)
```

```{r obukhovLength-check}
grade_result(
  pass_if(~identical(round(.result,1), round(12.02854,1)))
)
```

### Gradient-diffusion theory

Consider the following covariances:

* $\overline{w'c'}$
* $\overline{w'u'}$ and $\overline{w'v'}$
* $\overline{w'\theta'}$

```{r turbulentFluxes01, echo = FALSE}

terms <- c("Gases", "Momentum", "Heat")
rank_terms <- sortable::question_rank(
  "Sort the following terms in the order they appear above",
  learnr::answer(terms, correct = TRUE),
  learnr::answer(rev(terms), correct = FALSE, "Other direction!"),
  allow_retry = TRUE
)
rank_terms
```

Consider the following equations:

* $H_{se} = \rho_m c_p \overline{w'\theta'}$
* $E = \rho_m \overline{w'c'}$
* $\tau = -\rho_m \overline{w'u'}$

```{r turbulentFluxes02, echo = FALSE}

terms <- c("Sensible heat flux", "Water vapor flux", "Momentum flux")
rank_terms <- sortable::question_rank(
  "Sort the following terms in the order of the equations given above",
  learnr::answer(terms, correct = TRUE),
  learnr::answer(rev(terms), correct = FALSE, "Other direction!"),
  allow_retry = TRUE
)
rank_terms
```

Consider the following units:

* mol m^-1^ s^-2^
* W m^-2^
* mol m^-2^ s^-1^

```{r turbulentFluxes03, echo = FALSE}

terms <- c("Momentum flux", "Sensible heat flux", "Water vapor flux")
rank_terms <- sortable::question_rank(
  "Sort the following terms in the order of their units given above",
  learnr::answer(terms, correct = TRUE),
  learnr::answer(rev(terms), correct = FALSE, "Other direction!"),
  allow_retry = TRUE
)
rank_terms
```

### The logarithmic wind profile

Consider a canopy with an aerodynamic roughness length of 2 m, a displacement height of 10 m, and a friction velocity of 0.3 m s^-1^. Assuming a logarithmic wind profile, calculate the horizontal wind speed at 40 m above the ground.

```{r logWindProfile, exercise = TRUE, exercise.eval = FALSE}

```

```{r logWindProfile-solution}
# Eq. 13.15, p. 306) 

# inputs
uStar <- 0.3 # friction velocity in m/s (see Holton, page 268)
k <- 0.4 # von Karman constant
z0 <- 2 # aerodynamic roughness length in m
dH <- 10 # displacement heihgt in m
z <- 40 # height in m

u <- uStar / k * log((z - dH) / z0)
print(u)
```

```{r momentumFlux-check}
grade_result(
  pass_if(~identical(round(.result,1), round( 2.031038,1)))
)
```

Estimate the corresponding momentum flux expressed in mol m^-1^ s^-2^, assuming a surface pressure of 100 hPa and a mean air temperature of 293 K.

```{r momentumFlux, exercise = TRUE, exercise.eval = FALSE}

```

```{r momentumFlux-solution}
P <- 100000 # atmospheric pressure (Pa)
T <- 293 # air temperature (K)
R <- 8.314 # universal gas constant (J K–1 mol–1)

rho_m <- P * 100 / (R * T) # molar density of dry air (mol m^-3)

tau <- uStar^2 * rho_m # (mol m–1 s–2)

print(tau)
```

```{r logWindProfile-check}
grade_result(
  pass_if(~identical(round(.result,1), round(369.4578,1)))
)
```

### Monin-Obukhov Similarity Theory (MOST)

The impact of atmospheric static stability on the vertical wind profile can be estimated as:

$$ u(z) = \frac{u_*}{k} \ln \left( \frac{z - d_H}{z_0} - \psi_m(\zeta)\right) $$ 

Make a Figure that shows how windspeed vaies with $\psi_m(\zeta)$ (ranging from -2.0 to 1.0 (-)) for the following conditions:

* Height: 30 m
* Displacement height: 19 m
* Roughness length for momentum: 0.6 m
* Friction velocity: 0.4 m s^-1^

```{r psi, exercise = TRUE, exercise.eval = FALSE}

```

```{r psi-solution}
rm(list=ls())

# constants
k <- 0.4 # con Karman constant (-)

# inputs
z <- 30 # height (m)
dH <- 19 # displacement height (m)
z_0m <- 0.6 # aerodynamic roughness length
uStar <- 0.4 # friction velocity (m s-1)

psi_m <- seq(-2, 1, 0.1) # similarity function (-)

u <- uStar / k * (log((z-dH) / z_0m) - psi_m) # wind speed (m s-1)

my.xlab <- "$psi_m$ (-)"
my.ylab <- "Wind speed (m s$^{-1}$)"

my.xlab <- latex2exp::TeX(my.xlab)
my.ylab <- latex2exp::TeX(my.ylab)

par(mar = c(5,5,1,1))
plot(x = psi_m, y = u, type = "l", lty = 1,
        xlab = my.xlab, ylab = my.ylab, las = 1)
```

```{r windspeedAndPsi, echo=FALSE}
question_radio(
  "Considering the Figure you made above, how does the windspeed respond as the atmosphere becomes increasingly unstable?",
  answer("Windspeed increases", correct = TRUE),
  answer("Windspeed decreases"),
  answer("Windspeed remains unaffected"),
  allow_retry = TRUE
)
```

Make a Figure that shows how the wind profile is affected by atmospheric stability for the following conditions:

* Displacement height: 19 m
* Roughness length for momentum: 0.6 m
* Friction velocity: 0.4 m s^-1^
* Monin-Obukhov length: -1000, -50, and -10m

Again, the impact of atmospheric static stability on the vertical wind profile can be estimated as:

$$ u(z) = \frac{u_*}{k} \ln \left( \frac{z - d_H}{z_0} - \psi_m(\zeta)\right) $$ 

* For unstable conditions ($\zeta < 0$): 

$$\psi_m(\zeta) = 2 \ln \left[ (1+ \chi) /2 \right] + \ln \left[ (1+\chi^2) / 2 \right] - 2 \tan^{-1} \chi + \pi/2$$

* where $\chi = (1-16 \zeta)^{1/4}$

```{r MOST01, exercise = TRUE, exercise.eval = FALSE}

```

```{r MOST01-solution}
rm(list=ls())

# constants
k <- 0.4 # con Karman constant (-)

# inputs
dH <- 19 # displacement height (m)
z_0m <- 0.6 # aerodynamic roughness length
uStar <- 0.4 # friction velocity (m s-1)


L <- c(-1000, -50, -10) # Obukhov length (m)

z <- seq(20,50,1)  

my.fun <- function(z, L){

zeta <- (z - dH) / L 
chi <- (1 - 16 * zeta)^(1/4)

termI <- 2 * log((1 + chi)/2)
termII <- log((1 + chi^2)/2)
termIII <- (-2) * atan(chi)
termIV <- pi / 2

psi_m <- termI + termII + termIII + termIV

u <- uStar / k * (log((z-dH) / z_0m) - psi_m) # wind speed (m s-1)

}

data <- outer(z, L, my.fun)

my.xlab <- "Height (m)"
my.ylab <- "Wind speed (m s$^{-1}$)"

my.xlab <- latex2exp::TeX(my.xlab)
my.ylab <- latex2exp::TeX(my.ylab)

my.col <- viridis::plasma(n = ncol(data),  begin = 0, end = 0.75)

par(mar = c(5,5,1,1))
matplot(x = z, y = data, type = "l", lty = 1, col = my.col,
        xlab = my.xlab, ylab = my.ylab, las = 1)

legend.text <- paste("L = ", L, "m", sep = " ")
legend("topleft", legend = legend.text, pch = 16, col = my.col, bty = "n")
```

```{r windspeedProfile, echo=FALSE}
question_radio(
  "Considering the Figure you made above, how does the increase of windspeed with height respond as the atmosphere becomes increasingly unstable?",
  answer("Windspeed increases more strongly", correct = TRUE),
  answer("Windspeed increases less strongly"),
  answer("Windspeed remains unaffected"),
  allow_retry = TRUE
)
```

Make a Figure that shows how the potential temperature profile is affected by atmospheric stability for the following conditions:

* Displacement height: 19 m
* Roughness length for momentum: 0.6 m
* Roughness length for scalars: 0.135 x roughness length for momentum
* Potential temperature scale: -0.5 K
* Monin-Obukhov length: -1000, -50, and -10m

The impact of atmospheric static stability on the vertical wind profile can be estimated as:
$$ \theta(z)-\theta_s = \frac{\theta_*}{k} \ln \left( \frac{z - d_H}{z_{0h}} - \psi_h(\zeta)\right) $$ 

* For unstable conditions ($\zeta < 0$): 

$$\psi_h(\zeta) = 2 \ln \left[ (1+ \chi^2) /2 \right]$$

* where $\chi = (1-16 \zeta)^{1/4}$

```{r MOST02, exercise = TRUE, exercise.eval = FALSE}

```

```{r MOST02-solution}
rm(list=ls())

# constants
k <- 0.4 # con Karman constant (-)

# inputs
dH <- 19 # displacement height (m)
z_0m <- 0.6 # aerodynamic roughness length for momentum
z_0h <- 0.135 * z_0m # aerodynamic roughness length for potential temperature (m)
thetaStar <- -0.5 # Potential temperature scale (K)


L <- c(-1000, -50, -10) # Obukhov length (m)

z <- seq(20,50,1)  

my.fun <- function(z, L){

zeta <- (z - dH) / L 
chi <- (1 - 16 * zeta)^(1/4)

psi_h <- 2 * log((1 + chi^2)/2)

dTheta <- thetaStar / k * (log((z-dH) / z_0h) - psi_h) # wind speed (m s-1)

}

data <- outer(z, L, my.fun)

my.xlab <- "Height (m)"
my.ylab <- "$theta(z) - theta_s$ (K)"

my.xlab <- latex2exp::TeX(my.xlab)
my.ylab <- latex2exp::TeX(my.ylab)

my.col <- viridis::plasma(n = ncol(data),  begin = 0, end = 0.75)

par(mar = c(5,5,1,1))
matplot(x = z, y = data, type = "l", lty = 1, col = my.col,
        xlab = my.xlab, ylab = my.ylab, las = 1)

legend.text <- paste("L = ", L, "m", sep = " ")
legend("topright", legend = legend.text, pch = 16, col = my.col, bty = "n")

```

```{r potTempProfile, echo=FALSE}
question_radio(
  "Considering the Figure you made above, how does the decrease of potential temperatrue with height respond as the atmosphere becomes increasingly unstable?",
  answer("Potential temperature decreases more strongly", correct = TRUE),
  answer("Potential temperature decreases less strongly"),
  answer("Potential temperature remains unaffected"),
  allow_retry = TRUE
)
```

Make a figure that shows how the similarity functions $\psi_m(\zeta)$ and $\psi_h(\zeta)$ vary with $\zeta$ ranging from -2 to 1.

```{r aerodynCond, exercise = TRUE, exercise.eval = FALSE}

# You will need to make an if-else statement here, e.g. if zeta < 0, then a, else b:
zeta <- seq(-2,1,0.5)

a <- 1
b <- 2
psi <- ifelse(zeta < 0, a, b)
print(psi)

```

```{r aerodynCond-solution}
rm(list=ls())

zeta <- seq(-2,1,0.1)

# psi
chi <- (1 - 16 * zeta)^(1/4)
termI <- 2 * log((1 + chi)/2)
termII <- log((1 + chi^2)/2)
termIII <- (-2) * atan(chi)
termIV <- pi / 2

psi_m <- ifelse(zeta < 0, termI + termII + termIII + termIV, (-5) * zeta)
psi_h <- ifelse(zeta < 0, 2 * log((1 + chi^2)/2), (-5) * zeta)

my.xlab <- "$zeta = z/L$ (-)"
my.ylab <- "$psi(zeta)$ (-)"

my.xlab <- latex2exp::TeX(my.xlab)
my.ylab <- latex2exp::TeX(my.ylab)

my.col <- viridis::plasma(n = 2,  begin = 0, end = 0.75)

par(mar = c(5,5,1,1))
plot(zeta, psi_h, xlab = my.xlab, ylab = my.ylab, las = 1, type = "l", col = my.col[1])
lines(zeta, psi_m, col = my.col[2])

abline(v = 0, h = 0)

legend.text <- latex2exp::TeX(c("$psi_h(z)$", "$psi_m (z)$"))
legend("topright", legend = legend.text, pch = 16, col = my.col, bty = "n")

```

```{r psiVsZeta, echo=FALSE}
question_radio(
  "Considering the Figure you made above, how does static stability affect the psi functions and hence wind speed and potential temperature?",
  answer("Psi increases as atmospheric conditions become increasingly unstable", correct = TRUE),
  answer("Psi decreases as atmospheric conditions become increasingly unstable"),
  answer("Psi remains unaffected"),
  allow_retry = TRUE
)
```