## Soil fluxes

### Decomposition of soil organic matter

```{r carbonPoolSizes, echo = FALSE}

terms <- c(
  "Dissolved inorganic carbon in the ocean", 
  "Soils", 
  "Permafrost",
  "Fossil fuel reserves",
  "Atmosphere", 
  "Organic carbon in the ocean",
  "Vegetation")

rank_terms <- sortable::question_rank(
  "Sort the following carbon pools with respect to their size in decreasing order",
  learnr::answer(terms, correct = TRUE),
  learnr::answer(rev(terms), correct = FALSE, "Other direction!"),
  allow_retry = TRUE
)
rank_terms
```

```{r carbonFluxSizes, echo = FALSE}

terms <- c(
  "Fossil fuel emissions", 
  "Land uptake", 
  "Ocean uptake",
  "Land use change",
  "Volcanic emissions")

rank_terms <- sortable::question_rank(
  "Sort the following carbon fluxes with respect to their size in decreasing order",
  learnr::answer(terms, correct = TRUE),
  learnr::answer(rev(terms), correct = FALSE, "Other direction!"),
  allow_retry = TRUE
)
rank_terms
```

```{r mineralization, echo=FALSE}
question_radio(
  "The conversion of inorganic nutrients to organic compounds is called:",
  answer("Mineralization"),
  answer("Immobilization", correct = TRUE)
)
```

```{r respiration, echo=FALSE}
question("Which of the following sources of CO2 fluxes contribute to rhizospheric respiration? (Select all that apply)",
answer("Root respiration", correct = TRUE),
answer("Microbial rhizosphere respiration", correct = TRUE),
answer("Microbial SOM substrate respiration"),
  allow_retry = TRUE
)
```

```{r residenceTime, echo=FALSE}
question_radio(
  "Which of the two processes below do you expect to have a longer residence time?",
  answer("Rhizospheric respiration"),
  answer("Decomposition respiration", correct = TRUE)
)
```

```{r soilPools, echo = FALSE}

terms <- c(
  "disaccharides, simple sugars, amino acids, and short peptides", 
  "cellulose and hemicellulose", 
  "lignin and lignocelluloses")

rank_terms <- sortable::question_rank(
  "Sort the compounds listed below in the order of their corresponding pools: active pool, slow pool, passive pool",
  learnr::answer(terms, correct = TRUE),
  learnr::answer(rev(terms), correct = FALSE, "Other direction!"),
  allow_retry = TRUE
)
rank_terms
```

Consider a sample of litter that consists of proteins, cellulose/hemicellulose, and lignin, with turnover coefficients of $k_1 = 0.1$, $k_2 = 0.05$, and $k_3 = 0.025$ day^-1^, respectively. Make a figure that shows how the remaining litter fractions change within one year time.

```{r decay01, exercise = TRUE, exercise.eval = FALSE}

```

```{r, decay01-solution}

# inputs

k <- c(0.1, 0.05, 0.025) # turnover coefficients
t <- seq(0, 365, 1) # time in days

# Equation Eq. 16.9, p.392
my.fun <- function(t, k){

  frac <- exp(-k * t) 

}

data <- outer(t, k, my.fun)

my.xlab <- "Time (days)"
my.ylab <- "Remaining fraction (-)"

my.col <- viridis::plasma(n = ncol(data),  begin = 0, end = 0.75)

par(mar = c(5,5,1,1))
matplot(x = t, y = data, type = "l", lty = 1, col = my.col,
        xlab = my.xlab, ylab = my.ylab, las = 1)

legend.text <- paste("k = ", k, "(-)", sep = " ")
legend("topright", legend = legend.text, pch = 16, col = my.col, bty = "n")

```

Consider a sample of litter that consists of cellulose with turnover coefficients of $k = 0.02$. The decomposition of the cellulose can be estimated using a $Q_{10}$ function:

$$ \frac{dc}{dt} = -k \times f_{15}(Q_{10}) \times c$$
$$f_{15}(Q_{10}) = Q_{10}^{(0.1(Tc-15))}$$

Make a figure that shows how the remaining cellulose fraction changes within one year time for different annual mean temperatures (ranging from 0 to 30 degrees C) and a $Q_{10}$ value of 2.

```{r decay02, exercise = TRUE, exercise.eval = FALSE}

```



```{r, decay02-solution}

# inputs
k <- 0.02 # turnover coefficient (-)
Q_10 <- 2 # Q10 value (-)
Tc <- seq(0, 30, 5) # temperature (C)
t <- seq(0, 365, 1) # time in days

# Equation Eq. 16.9, p.392
my.fun <- function(t, Tc){

  f_15 <- Q_10^(0.1 * (Tc - 15))
  frac <- exp(-k * f_15 * t) 

}

data <- outer(t, Tc, my.fun)

my.xlab <- "Time (days)"
my.ylab <- "Remaining fraction (-)"

my.col <- viridis::plasma(n = ncol(data),  begin = 0, end = 0.75)

par(mar = c(5,5,1,1))
matplot(x = t, y = data, type = "l", lty = 1, col = my.col,
        xlab = my.xlab, ylab = my.ylab, las = 1)

legend.text <- paste("T = ", Tc, "(C)", sep = " ")
legend("topright", legend = legend.text, pch = 16, col = my.col, bty = "n")

```

Turn to your neighbor and discuss the implications of your Figure above for climate change.

### Methane and Nitrogen oxides emissions from soils

Turn to your neighbor and discuss why methanogenesis would not prevail under aerobic conditions.


```{tikz, Ncycle, fig.ext = 'png', cache=TRUE}
\usetikzlibrary{arrows}
\begin{tikzpicture}[node distance=1.0cm, scale = 1.0]
\node (12) at (2, 2) {N$_2$ (or N$_2$O)};
\node (3) at (4, 0) {NO$_2^-$/NO$_3^{-}$};
\node (6) at (2, -2) {\color{white}, Organic N};
\node (9) at (0, 0) {\color{white}, NH$_3$/NH$_4^+$};
\node (6b) at (1.8, -2) {Organic N};
\node (9b) at (-0.2, 0) {NH$_3$/NH$_4^+$};
% arrows
\draw [->] (12) -- (9) node[midway, anchor=south]{1};
\draw [->] (9) -- (6) node[midway, anchor=south]{2};
\draw [->] (9) -- (3) node[midway, anchor=south]{3};
\draw [->] (3) -- (6) node[midway, anchor=south]{4};
\draw [->] (6b) -- (9b) node[midway, anchor=north]{5};
\draw [->] (3) -- (12) node[midway, anchor=south]{6};
\end{tikzpicture}
```

```{r NcycleSort, echo = FALSE}

terms <- c(
"Nitrogen fixation",
"Ammonia assimilation",
"Nitrification",
"Assimilatory nitrate reduction",
"Ammonification",
"Denitrification")

rank_terms <- sortable::question_rank(
  "Sort the following terms in the order of arrows depicted in the Figure above",
  learnr::answer(terms, correct = TRUE),
  learnr::answer(rev(terms), correct = FALSE, "Other direction!"),
  allow_retry = TRUE
)
rank_terms
```

```{r NcycleEnergy, echo=FALSE}
question("Which of the following processes represent a source of energy? (Select all that apply)",
answer("Nitrogen fixation"),
answer("Ammonia assimilation"),
answer("Nitrification", correct = TRUE),
answer("Assimilatory nitrate reduction"),
answer("Ammonification", correct = TRUE),
answer("Denitrification", correct = TRUE),
  allow_retry = TRUE
)

```{r NcycleAerobic, echo=FALSE}
question("Which of the following processes require aerobic conditions? (Select all that apply)",
answer("Nitrogen fixation"),
answer("Ammonia assimilation"),
answer("Nitrification", correct = TRUE),
answer("Assimilatory nitrate reduction"),
answer("Ammonification", correct = TRUE),
answer("Denitrification"),
  allow_retry = TRUE
)

```{r NcycleAnaerobic, echo=FALSE}
question("Which of the following processes require anaerobic conditions? (Select all that apply)",
answer("Nitrogen fixation", correct = TRUE),
answer("Ammonia assimilation"),
answer("Nitrification"),
answer("Assimilatory nitrate reduction"),
answer("Ammonification"),
answer("Denitrification", correct = TRUE),
  allow_retry = TRUE
)

