
## Programming in R
In this course you will be using the programming language R. Today you will learn the R basics you need to master the course. The main elements are: 

* Create a new project
* Operators: ```+, -, *, /, ^```
* R objects (vector, data frame, matrix, array, list)
* Subset objects
* Control structure: loops, if-else statements
* Generating data: sequence and random data
* Defining a function
* Plotting data

Let's start with setting up your computer. R is a command line driven program. The user enters commands at the prompt and each command is executed one at a time. RStudio is a Graphic User Interface for R. For this course, you will first need to install R and then RStudio. If you already have R or RStudio installed, please install them again to ensure you have the most recent version.

* Install R from [here](https://posit.co/download/rstudio-desktop)
* Install RStudio from [here](https://posit.co/download/rstudio-desktop)
* Once R and RStudio are installed, open RStudio and learn the basic R syntax following the exercises below.

### Create a new Project
* Open RStudio
* Click on ```File``` in the upper-left corner
* From the drop-down list choose ```New Project ...```, ```New Directory```, ```Browse...``` and choose a directory of your project. For instance, in my case this will be ```C:\Users\seile\Documents\queensu\teaching\classes\ENSC430\tutorials\ENSC430```. Important: make sure your directory does not include white space or special characters.
* Go to your new directory and create a new folder called ```renv```. This is where you will store your packages, as explained next.

### Operators

* In R, you use the following syntax to assign a value to a variable.

```{r, eval=FALSE, echo=TRUE}
x <- 1
print(x)
```

* The variable x used above is an example of an R object. The ```print``` statement prints the value of the object x.
* Once you assign values to R objects, you can use the following arithmetic operators to perform calculations.

```{r, eval=FALSE, echo=TRUE}
a <- 2
b <- 3

# Addition
c <- a+b
print(c)

# Subtraction
c <- a-b
print(c)

# Multiplication
c <- a*b
print(c)

# Division
c <- a/b
print(c)

# Exponentiation
c <- a^b
print(c)
```
* Note that anything after the ```#``` symbol is ignored.

```{r, eval=FALSE, echo=TRUE}
a <- 2
b <- 3
c <- a + b
# c <- (a + b)^2
print(c)
```
* This way you can add comments into your code that to explain what you are doing

### Objects

* All objects have two intrinsic attributes: mode and length. 
* The mode is the basic type of the elements of the object. 
* There are four main modes: numeric, character, complex, and logical (FALSE or TRUE). 

```{r, eval=FALSE, echo=TRUE}
a <- 5
class(a)

a <- "apple"
class(a)

a <- 1 + 2i
class(a)

a <- TRUE
class(a)
```
* The length of an object is the number of elements of the object.

```{r, eval=FALSE, echo=TRUE}
a <- 2
length(a)

a <- c(2,7,5)
length(a)
```
* Next to mode and length, there are different types of objects, namely vector, data frame, matrix, array, and list

```{r, eval=FALSE, echo=TRUE}
# Vector with one element
a <- 2
print(a)

# Vector with four elements
b <- c(7,5,2,9)
print(b)

# Data frame
x <- c(3,7,2,4)
y <- c(42,67,21,33)
df <- data.frame(x, y)
print(df)

# Matrix: in R, a matrix has 2 dimensions only)
x <- c(1,2,3,4,5,6,7,8,9,10)
a <- matrix(x, nrow = 2)
print(a)

# Array: an array has more than 2 dimensions
x <- c(1,2,3,4,5,6,7,8,9)
a <- array(data = x, dim = c(3,3,3))
print(a)

# List
a <- c(3,2,5,6)
b <- c(TRUE, FALSE)
c <- c("apples", "pears", "plums")
d <- 5

l <- list(a,b,c,d)
print(l)
```

* Let's learn how to subset vectors, data frames, matrices, arrays, and lists

* Example: Vector
```{r, eval=FALSE, echo=TRUE}
# Create a vector
vector_data <- c(10, 20, 30, 40, 50)

# Print the vector
print("Full Vector:")
print(vector_data)

# Subsetting by Index
third_element <- vector_data[3]
print("Third element:")
print(third_element)

# Subsetting by Multiple Indices
subset_indices <- vector_data[c(1, 3, 5)]
print("Subset (First, Third, and Fifth elements):")
print(subset_indices)

# Subsetting by Logical Condition
subset_condition <- vector_data[vector_data > 25]
print("Subset (Elements greater than 25):")
print(subset_condition)
```

* Example: data frame
```{r, eval=FALSE, echo=TRUE}
# Create a data frame
data <- data.frame(
  Name = c("Alice", "Bob", "Charlie", "David", "Eve"),
  Age = c(23, 25, 30, 22, 29),
  Gender = c("F", "M", "M", "M", "F"),
  Score = c(85, 90, 75, 88, 92)
)

# Print the data frame
print("Full Data Frame:")
print(data)

# Subsetting by Columns
subset_columns <- data[, c("Name", "Score")]
print("Subset by Columns (Name and Score):")
print(subset_columns)

# Subsetting by Rows and Columns
subset_rows_cols <- data[1:3, c("Name", "Age")]
print("Subset by Rows and Columns (First 3 rows, Name and Age):")
print(subset_rows_cols)

# Subsetting by Condition
subset_condition <- subset(data, Age > 25)
print("Subset by Condition (Age > 25):")
print(subset_condition)
```

* Example: matrix
```{r, eval=FALSE, echo=TRUE}
# Create a matrix
matrix_data <- matrix(
  1:12,        # Data
  nrow = 4,    # Number of rows
  ncol = 3,    # Number of columns
  byrow = TRUE # Fill the matrix by rows
)

# Print the matrix
print("Full Matrix:")
print(matrix_data)

# Subsetting by Rows and Columns
subset_rows_cols <- matrix_data[1:2, 1:2]
print("Subset (First 2 rows, First 2 columns):")
print(subset_rows_cols)

# Subsetting by Specific Rows
subset_specific_rows <- matrix_data[c(1, 3), ]
print("Subset (First and Third rows):")
print(subset_specific_rows)

# Subsetting by Specific Columns
subset_specific_cols <- matrix_data[, c(2, 3)]
print("Subset (Second and Third columns):")
print(subset_specific_cols)
```

* Example: array
```{r, eval=FALSE, echo=TRUE}

# Create an array
array_data <- array(
  1:24,         # Data
  dim = c(3, 4, 2) # Dimensions (3x4x2 array)
)

# Print the array
print("Full Array:")
print(array_data)

# Subsetting by Indices
element <- array_data[1, 2, 1]
print("Element at (1, 2, 1):")
print(element)

# Subsetting by Slices
slice <- array_data[, , 1]
print("First slice (all elements in the first third dimension):")
print(slice)

# Subsetting by Specific Rows and Columns in a Specific Dimension
subset_array <- array_data[1, , 2]
print("First row, all columns, in the second slice:")
print(subset_array)
```

* Example: list
```{r, eval=FALSE, echo=TRUE}
# Create a list
list_data <- list(
  Name = "Alice",
  Age = 25,
  Scores = c(85, 90, 88),
  Address = list(
    Street = "123 Main St",
    City = "Springfield",
    Zip = "12345"
  )
)

# Print the list
print("Full List:")
print(list_data)

# Subsetting by Element Name
name <- list_data$Name
print("Name element:")
print(name)

# Subsetting by Index
age <- list_data[[2]]
print("Second element (Age):")
print(age)

# Subsetting a Nested List
city <- list_data$Address$City
print("City element in Address:")
print(city)
```

### Control structure
* Control structures determine the order in which statements are executed in a program. Two important examples are loops and conditional statements.
* Loop example
```{r, eval=FALSE, echo=TRUE}
# For loop to print numbers from 1 to 5
for (i in 1:5) {
  print(i)
}
```

* Conditional statement example
```{r, eval=FALSE, echo=TRUE}

# Define a number
number <- -3

# If-else statement to check the sign of the number
if (number > 0) {
  print("The number is positive.")
} else if (number < 0) {
  print("The number is negative.")
} else {
  print("The number is zero.")
}

```

### Generating data
* Many times you need to generate data, e.g. a sequence that ranges from 1 to 10. Instead of writing out each number, there are functions that do that for you

```{r, eval=FALSE, echo=TRUE}
# Sequence from 1 to 10 with step size 1
x <- 1:10
print(x)

# Sequence from 1 to 10 with step size 0.5
x <- seq(1, 10, 0.5)
print(x)

# Vector that repeats the value one for ten times
x <- rep(1, 10)
print(x)

# Random sequences with uniform distribution between zero and one
x <- runif(n = 5, min = 0, max = 1)
print(x)
```

### Defining a function
* Often you need to execute the same operation many times. To do this efficiently you can define a function and then apply this function whenever you need it

```{r, eval=FALSE, echo=TRUE}
# Define a function to add two numbers
add_numbers <- function(a, b) {
  # Calculate the sum
  sum <- a + b
  # Return the sum
  return(sum)
}

# Print the function definition
print("Function Definition:")
print(add_numbers)

# Call the function with arguments 3 and 5
result <- add_numbers(3, 5)

# Print the result
print("Result of add_numbers(3, 5):")
print(result)
```

### Plotting data
```{r, eval=FALSE, echo=TRUE}
x <- seq(1:10)
y <- x^2
plot(x, y)
```

* Add x and y labels using ```xlab``` and ```ylab```
```{r, eval=FALSE, echo=TRUE}
x <- seq(1:10)
y <- x^2
plot(x, y, 
     xlab = "ENSC430 lectures attended", 
     ylab = "Academic Performance")
```

* Make it a line plot using ```type = "l"```
```{r, eval=FALSE, echo=TRUE}
x <- seq(1:10)
y <- x^2
plot(x, y, type = "l")
```

* Add more lines using ```lines()```
```{r, eval=FALSE, echo=TRUE}
x <- seq(1:10)
y1 <- x^2
y2 <- x^3
plot(x, y1, type = "l")
lines(x, y2)
```

* Make each line a different color using ```col```
```{r, eval=FALSE, echo=TRUE}
x <- seq(1:10)
y1 <- x^2
y2 <- x^3
plot(x, y1, type = "l", col = "red")
lines(x, y2, col = "blue")
```

* Adjust x and y axis using ```xlim``` and ```ylim```
```{r, eval=FALSE, echo=TRUE}
x <- seq(1:10)
y1 <- x^2
y2 <- x^3
plot(x, y1, type = "l", 
     xlim = c(-5,10), ylim = c(-20, 100))
lines(x, y2)
```

* Making square axis using ```par(pty="s")``` and ```asp = 1```
```{r, eval=FALSE, echo=TRUE}
x <- seq(1:10)
y1 <- x^2
y2 <- x^3
par(pty="s")
plot(x, y1, type = "l", asp = 1)
lines(x, y2)
```

* You will be plotting equations as a way to understand the dependence of two variables and the role of parameters
* These plotting exercises require that you plot the same function multiple times with different parameter values
* This can be done conveniently by using the functions ```outer``` and ```matplot```
* Below an example where you plot ```y``` as a function of ```x``` for different values of parameter ```p```

```{r, eval = TRUE, echo = TRUE}
# Parameter: give it sequence of values
p <- seq(1, 10, 1)

# independent variable
x <- seq(1,100,1)

# Define a function
my.fun <- function(x, p){
y <- x^(2 + p/100)
  return(y)
}

# Calculate all values of y for a given x and p:
data <- outer(x, p, my.fun)

# Have a look at data: Each column give you the value of y for a given x and p
head(data)

my.xlab <- "x"
my.ylab <- "y"

# Give each line a different color
my.col <- rainbow(n = ncol(data))

# Plot all lines
matplot(x = x, y = data, type = "l", lty = 1, col = my.col,
        xlab = my.xlab, ylab = my.ylab)

legend.text <- paste("p =", p, sep = " ")
legend("left", legend = legend.text, pch = 16, col = my.col, bty = "n")
```

Many times we need to add fancy text that includes Greek letters, exponents, and subscripts in Figures. That can be done using the LaTeX language, whihc is made available in R through the ```latex``` package. Text written in LaTeX is enclosed by dollar signs. Here some examples:

* Greek letters: ```$\alpha$``` yields $\alpha$

* Exponents: ```$m^{-2}$``` yields $m^{-2}$

* Subscripts: ```$x_i``` yields $x_i$

### Review
Congratulations, you made it to the end of this tutorial. Let's list the key concepts you have learned today:

* Operators: ```+, -, *, /, ^```
* Objects: vector ```c()```, data frame ```data.frame()```, matrix ```matrix()```, array ```array()```, list ```list()```, 
* Subset objects
* Control structure: loops, if-else statements
* Generating data: sequence and random data
* Defining a function
* Plotting data
* Plotting a function with varying parameter values

### Exercise (1)

Write a piece of code that includes the following elements:

* Create an R object that is either a data frame, matrix, array, or list
* Define a function
* Apply the function to a subset of the R object using either a loop or a conditional statement
* Plot the data


### Exercise (2)

* Consider the following equation: $y = \exp(x) + b$
* Plot this function for different values of $b$ using the ```matplot``` function
